worldmap

### 2021-05-13 analysis/05_model_metrics.Rmd
Reorganize

#get metric data
metrics <- pins::pin_get("rarefied-metrics", board = "github") %>%
  # pivot_longer(., cols = setdiff(everything(), one_of("rarefyID", "STUDY_ID", "SamplePool", "SampleN", "num_years", "duration", "startYear", "endYear", "YEAR", "cell", "rarefied", "type", "realm", "climate", "habitat", "biome_map", "taxa", "organisms", "cent_lat", "cent_long", "abundance_type", "BROAD_TYPE", "taxa_mod", "climate_mod")), names_to = "metric", values_to = "value") %>%
  rename_with(tolower, .cols = setdiff(everything(), one_of("rarefyID")))  %>%
  filter(!is.na(value)) %>%
  mutate(logvalue = log(value + 1)) %>% # create a column for log transformation that adds the absolute value integer closest to the most negative value to create all positive values
  left_join(meta, c("rarefyID" = "rarefyid"))
```

@@

How many years are missing FD metrics? Only a handful! Let's exclude them so we can get an accurate view of the number of years for which we have data for each timeseries                                                                                                     
```{r}
# no_fd_years <- metrics %>%
#   filter(metric %in% c("FRic", "FEve", "FDiv", "FDis"), is.na(value))
# 
# dim(no_fd_years)
# 
# metrics <- setdiff(metrics, no_fd_years)
```

@@

Fit GLM to overall metric trend

@@

```{r}
#find rarefyID's that have inf value for some metrics and exclude
cont_inf <- metrics %>% filter(is.infinite(value), !metric %in% c("qual.FRic",  "SES_qual.FRic")) %>% pull(rarefyID)

# model for overall trend (random effect w/varying intercept)
# metric_mean_fit <- model_data %>%
#   filter(metric %in%  c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(log(value + 1) ~ year_scaled + (1|study_id/rarefyID), data = .x, REML = FALSE,
#                   control = lmerControl(optCtrl = list(maxit = 1e9,
#                                                     maxfun = 1e9,
#                                                     xtol_abs = 1e-11,
#                                                     ftol_abs = 1e-11,
#                                                     maxeval = 1e9)))),
#          null_model = map(data, ~lmer(logvalue ~ (1|study_id/rarefyID), data = .x, REML = FALSE,
#                   control = lmerControl(optCtrl = list(maxit = 1e9,
#                                                     maxfun = 1e9,
#                                                     xtol_abs = 1e-11,
#                                                     ftol_abs = 1e-11,
#                                                     maxeval = 1e9)))),
#            
#            test = map2(model, null_model, anova),
#            test_res = map(test, tidy),
#            coef = map(model, tidy)
#            ) %>% 
#   unnest(cols = c(test_res)) %>%
#   select(metric, model, coef, p.value) %>%
#   filter(!is.na(p.value)) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))

>>>>

#get metric data
metrics <- pins::pin_get("rarefied-metrics", board = "github") %>%
  # pivot_longer(., cols = setdiff(everything(), one_of("rarefyID", "STUDY_ID", "SamplePool", "SampleN", "num_years", "duration", "startYear", "endYear", "YEAR", "cell", "rarefied", "type", "realm", "climate", "habitat", "biome_map", "taxa", "organisms", "cent_lat", "cent_long", "abundance_type", "BROAD_TYPE", "taxa_mod", "climate_mod")), names_to = "metric", values_to = "value") %>%
  rename_with(tolower, .cols = setdiff(everything(), one_of("rarefyID")))  %>%
  filter(!is.na(value)) %>%
  mutate(logvalue = log(value + 1)) %>% # create a column for log transformation that adds the absolute value integer closest to the most negative value to create all positive values
  left_join(meta, c("rarefyID" = "rarefyid"))

## Now the data that actually makes it into the model ##

#find rarefyID's that have inf value for some metrics and exclude
cont_inf <- metrics %>% filter(is.infinite(value), !metric %in% c("qual.FRic",  "SES_qual.FRic")) %>% pull(rarefyID)

pins::pin(model_data, board = "github")
```
Add human footprint data
```{r}
hfp_data <- pins::pin_get("hfp-data", board = "github")

model_data_hfp <- model_data %>% 
  #dplyr::select(rarefyID, year) %>% 
  #distinct() %>%
  mutate(hfp_year = case_when(
    year < 2001 ~ 1993,
    TRUE ~ 2009
  ), 
  hfp_rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
  hfp_rarefyID = str_replace_all(hfp_rarefyID, pattern = "_mamm", "")) %>%
  left_join(hfp_data %>% dplyr::select(-STUDY_ID) %>% mutate(year = as.numeric(year)), by = c("hfp_year" = "year", "hfp_rarefyID" = "rarefyID")) %>%
  dplyr::select(-hfp_year)

```

@@

@@

####Fit GLM to overall metric trend

@@

```{r}
# model for overall trend (random effect w/varying intercept)
# metric_mean_fit <- model_data %>%
#   filter(metric %in%  c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(log(value + 1) ~ year_scaled + (1|study_id/rarefyID), data = .x, REML = FALSE,
#                   control = lmerControl(optCtrl = list(maxit = 1e9,
#                                                     maxfun = 1e9,
#                                                     xtol_abs = 1e-11,
#                                                     ftol_abs = 1e-11,
#                                                     maxeval = 1e9)))),
#          null_model = map(data, ~lmer(logvalue ~ (1|study_id/rarefyID), data = .x, REML = FALSE,
#                   control = lmerControl(optCtrl = list(maxit = 1e9,
#                                                     maxfun = 1e9,
#                                                     xtol_abs = 1e-11,
#                                                     ftol_abs = 1e-11,
#                                                     maxeval = 1e9)))),
#            
#            test = map2(model, null_model, anova),
#            test_res = map(test, tidy),
#            coef = map(model, tidy)
#            ) %>% 
#   unnest(cols = c(test_res)) %>%
#   select(metric, model, coef, p.value) %>%
#   filter(!is.na(p.value)) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))

### 2021-05-13 analysis/visualization.Rmd
Reorganize

---
title: "Visualization"
author: "Kari Norman"
date: "9/18/2019"
output: html_document
---

@@

```{r}
library(tidyverse)
```

```{r}
metrics <- pins::pin_get("rarefied-medians", board = "github") %>%
  pivot_longer(., cols = setdiff(everything(), one_of("rarefyID", "STUDY_ID", "SamplePool", "SampleN", "num_years", "duration", "startYear", "endYear", "YEAR", "cell", "rarefied", "type", "realm", "climate", "habitat", "biome_map", "taxa", "organisms", "cent_lat", "cent_long", "abundance_type", "BROAD_TYPE", "cYEAR", "taxa_mod", "climate_mod")), names_to = "metric", values_to = "value") %>%
  rename_with(tolower, .cols = setdiff(everything(), one_of("rarefyID"))) 
```

```{r}
usa <- map_data("usa")

plt <- ggplot() +
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
  coord_fixed(1.3)

plt +
  geom_point(data = biotime_data %>% filter(taxa == "Fish"), aes(x = longitude, y = latitude), color = "black", size = .25)

```

#Visualizing FD trends
```{r}
#load(here::here("data", "fd_table.rda"))
fd_table <- pins::pin_get("fd-table", board = "github")

fric_plot <- fd_table %>%
  filter(metric == "FRic") %>%
  ggplot(aes(
    x = year,
    y = value,
    colour = as.factor(plot)
  )) +
  geom_line() +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  facet_wrap( ~ as.factor(plot_id), scales = "free") +
  theme(legend.position = "none") +
  ylab("Fric")

ggsave("fric_trends.png", fric_plot)
```

Reproduce Dornelas figure
```{r}
#get climate classification
metadata <- read_csv(system.file("extdata", "biotime/biotime_metadata.csv", package = "biodivTS"))

fric_trend <- fd_table %>%
  filter(metric == "FRic") %>%
  left_join(metadata %>% select(study_id, climate)) %>%
  ggplot(aes(x = year, y = log(value))) +
  # geom_smooth(aes(color = climate, group = plot_id), method = "glm", se = FALSE) +
  # geom_point(aes(color = climate, group = plot_id)) +
  geom_smooth(aes(group = plot_id), color = "grey", method = "glm", se = FALSE) +
  geom_point(aes(group = plot_id) ,color = "grey") +
  geom_smooth(color = "black", method = "glm", se = FALSE) +
  theme(legend.position = "none") 
ggsave("fric_trends.png", fric_plot)
```

metric plotting function
```{r}
library(lme4)

plot_metric <- function(metric_name, ylabel){
  data <- metrics %>%
    filter(metric == metric_name, !is.na(value), !is.infinite(value)) %>%
    mutate(logvalue = log(value))

  mod <- lmer(logvalue ~ year + (1|rarefyID), data = data)
  coefs <- summary(mod)$coefficients

  print(metric_name) 
  print(summary(mod))
  confint(mod)

  pred <- data.frame(year = min(data$year):max(data$year)) %>%
    mutate(pred = coefs[1,1] + year * coefs[2,1])

  trend_plot <- data %>%
    left_join(pred, by = "year") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = logvalue)) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey") +
    geom_smooth(aes(group = rarefyID), color = "darkgrey", method = "glm", se = FALSE) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme(legend.position = "none") +
    theme_classic() +
    ylab(ylabel)

  ggsave(paste0(metric_name, "_trend.png"), trend_plot)
  return(trend_plot)
}

plot_metric("FRic", "log(FD)")

plot_map <- tibble(metric_name = c("FRic", "FEve", "FDiv", "FDis"), 
           ylabel = c("log(FD)", "log(Evenness)", "log(Divergence)", "log(Dispersion)"))
pmap(plot_map, plot_metric)
```

get model info
```{r}
get_metric_conint <- function(metric_name){
 data <- metrics %>%
    filter(metric == metric_name) %>%
    mutate(rarefyID = as.factor(rarefyID), logValue = log(value))

  mod <- lmer(logValue ~ year + (1|rarefyID), data = data)
  coefs <- summary(mod)$coefficients

  confint(mod)
  # as_tibble(confint(mod)) %>%
  #   mutate(metric = metric_name)
}

map(c("FRic", "FEve", "FDiv", "FDis"), get_metric_conint)
```

>>>>

---
title: "worldmap"
output: html_document
---

@@

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
library(ggmap)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(mapproj)
library(stringr)
```

```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")
country_data <- pins::pin_get("country-data-bind", board = "github") %>%
  mutate(country = str_remove(country, "Commonwealth of The "),
         country = str_remove(country, "Kingdom of "),
         country = str_remove(country, "Republic of "),
         country = str_remove(country, "Kingdom of "),
         country = str_remove(country, "Federative Republic of "),
         country = str_remove(country, "Federative "),) %>%
  mutate(country = case_when(
    country == "United States" ~ "USA",
    country == "Russian Federation" ~ "Russia",
    country == "United Great Britain and Northern Ireland" ~ "UK",
    country == "Czechia" ~ "Czech Republic",
    TRUE ~ country
  )) %>% 
  count(country, code)

mapdata <- map_data("world") %>%
  left_join(country_data, by = c("region" = "country"))

mapdata %>%
  ggplot(aes(long, lat)) +
  geom_polygon(mapping = aes(group = group, fill = n)) +
  ggthemes::theme_map() +
  #scale_fill_continuous(color = "#32A251")
  scale_fill_gradient(low = "white", high = "#39737c")

ggsave("study_map.jpeg")
```

```{r}
#get rarefyid centers
included_ids <- pins::pin_get("model-data", board = "github") %>%
  select(rarefyID, taxa, climate, realm) %>%
  distinct() %>% 
  mutate(rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
  rarefyID = str_replace_all(rarefyID, pattern = "_mamm", ""))

study_data <- pins::pin_get("rarefyID-cell-centre", board = "github") %>%
  inner_join(included_ids)

map_data("world") %>%
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "grey") + 
  coord_map(xlim=c(-180,180), ylim=c(-57, 90)) +
  geom_point(data = study_data, aes(rarefyID_x, rarefyID_y, color = taxa), alpha = 0.5, inherit.aes = FALSE) +
  ggthemes::theme_map()
```
